Course Project 1
------------------------------------------

### Loading and preprocessing the data
```{r, include=FALSE}
cp1 <- read.csv("activity.csv", header = TRUE)
cp1$date <- as.Date(cp1$date)
library(dplyr)
```

```{r, echo=FALSE}
head(cp1,10)
```


### What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day

```{r, include=FALSE}
total_steps <- cp1 %>% group_by(date) %>% summarize(total_steps = sum(steps, na.rm = TRUE))
```

```{r, echo=FALSE}
head(total_steps, 10)
```

2. Histogram

```{r, echo=FALSE}
hist(total_steps$total_steps, breaks = 10, xlab ="Steps per day", main="Steps per day Histogram")
```


3. Calculate mean and median

```{r, echo=FALSE}
mean(total_steps$total_steps)
median(total_steps$total_steps)
```


### What is the average daily activity pattern?

1. Series plot of the 5-minute interval and the average number of steps taken

```{r, include=FALSE}
total_interval <- cp1 %>% group_by(interval) %>% summarize(total_interval = mean(steps, na.rm = TRUE))
```

```{r, echo=FALSE}
plot(total_interval$interval, total_interval$total_interval, type = "l", xlab = "5-min interval", ylab="Average steps", main="Average daily steps per interval" )
```


2. 5-minutes interval with the maximum number of steps

```{r, include=FALSE}
maxsteps <- max(total_interval$total_interval)
max_interval <- subset(total_interval, total_interval  >= maxsteps)
print2 <- max_interval[1,1]
```

```{r, echo=FALSE}
print(print2)
```


### Imputing missing values

1. Total Nas

```{r, echo=FALSE}
sum(is.na(cp1))
```


2/3.Substitute NA by mean and create the dataset

```{r, include=FALSE}
cp1nNa <- cp1 %>% mutate_all(~ifelse(is.na(.x), mean(.x, na.rm = TRUE), .x))
cp1nNa$date <- as.Date(cp1nNa$date, origin="1970-01-01")
```

```{r, echo=FALSE}
head(cp1nNa,10)
```

4. Make the histogram an calculate mean and median

```{r, include=FALSE}
total_stepsnNa <- cp1nNa %>% group_by(date) %>% summarize(total_stepsnNa = sum(steps, na.rm = TRUE))
```

```{r, echo=FALSE}
hist(total_stepsnNa$total_stepsnNa, breaks = 10, xlab ="Steps per day", main="Steps per day Histogram")
```

```{r, echo=FALSE}
mean(total_stepsnNa$total_stepsnNa)
median(total_stepsnNa$total_stepsnNa)
```


### Are there any differences in activity patterns between weekdays and weeends?

1. Create new vector with variables weekday and weekend

```{r, include=FALSE}
cp1$weekday <- weekdays(cp1$date)
cp1$weekend <- grepl("S.+", cp1$weekday)
```

```{r, echo=FALSE}
head(cp1,10)
```

2. Plot comparing weekdays and weekends

```{r, echo=FALSE}
par(mfrow = c(2, 1))
weekend <- filter(cp1, weekend == TRUE)
weekday <- filter(cp1, weekend == FALSE)

total_interval_weekday <- weekday %>% group_by(interval) %>% summarize(total_interval_weekday = mean(steps, na.rm = TRUE))
plot(total_interval_weekday$interval, total_interval_weekday$total_interval_weekday, type = "l", xlab = "5-min interval", ylab="Average steps", main="Average daily steps per interval" )


total_interval_weekend <- weekend %>% group_by(interval) %>% summarize(total_interval_weekend = mean(steps, na.rm = TRUE))
plot(total_interval_weekend$interval, total_interval_weekend$total_interval_weekend, type = "l", xlab = "5-min interval", ylab="Average steps", main="Average daily steps per interval" )
```


